Preregistration Template for Modelling in Ecology and Conservation

### 2023-10-04 _analysis/preregistration_template.Rmd
Delete 'about the template' from previous draft of fully described preregistration template.

# About this template

This template has been collated from a combination of literature reviews of the 'Good Modeling Practice' and other modeling literature, as well as a collaborative workshop at DELWP/ARI. Chris Jones has kindly volunteered one component of the VEFMAP project investigating vegetation responses to environmental flows as a testing ground for the use of preregistration templates in applied ecological modeling research.

The template is largely structured around 6 "phases" in the process of developing, evaluating and analysing a model or models in ecological applications: `Problem Formulation`, `Define Conceptual Model`, `Formalise and Specify Model`,  `Model Validation and Evaluation`, `Model Analysis`, `Model Application`. These have been defined and described to some extent throughout the template. This text (usually italicised) will eventually be converted to 'help-text', in the form-of mouse-over pop-ups, or else prompt-text within a text-box. Other parts of this text will be pulled into a supporting diagram describing the modelling workflow stages.

## Adaptive Preregistration

The following is an excerpt from my manuscript as at 21 September 2020. It describes the procedure for how this template is meant to be used within a modeling context. I hope to write this up as a guide (including screen-shots) that will accompany the template.

Adaptive Preregistration for model-based research using GitHub
We propose leveraging both the version-control and the project management features of GitHub <link to GitHub> as a tool for implementing adaptive preregistration <link to box>. When a document is version-controlled using git and GitHub, each version of a document is time-stamped, and is assigned its own unique document identifier, or commit-hash. Because GitHub documents are ‘time-stamped’ the genesis of the preregistration from one version to the next is apparent. Moreover, using the diff view of a document on GitHub, exactly how a document has changed between versions is clear - additions are coloured green, deletions are coloured red, and changes within a line are highlighted <insert picture>. Since the preregistration document would exist within the project repository containing the data and modelling and analysis code itself, any results from analyses that cause the analysis plan to change can be linked and referenced in the updated version of the plan — thus facilitating the process of model checking. By providing a method for contrasting the actual analysis undertaken against the registered analysis, this is a mechanism for allowing reviewers to properly evaluate the preregistration against the reported analysis. It thus provides a way of marking preregistered parts of a report from non-preregistered parts of the analysis.

How it works….

GitHub is used to track amendments to the preregistration — each time an additional question is answered, or an alteration is made to the preregistration, the change should be committed to the repository. Semantic versioning is used to document each version of the preregistration. Semantic versioning is an open-source software engineering practice used to document and and communicate the development of software, in a way that allows others to depend on it  . We have created our own semantic versioning scheme to be used to specifically document the incremental version of the preregistration, in tandem with GitHub’s tag and release feature, each interim preregistration is tagged and released with its semantic version number.

All questions that can be answered should be completed before any data analysis or modelling proceeds. The preregistration should be completed until a) no further detail can be supplied or b) no additional template item can be answered until preliminary investigations or analyses are undertaken (Fig X, steps 1 - 3).

Shifting from planning to execution

Before shifting from planning to the execution of the interim preregistration, each complete response in the interim preregistration should have a corresponding GitHub issue created (Figure X, step 3, document). <define GitHub issue, https://guides.github.com/features/issues/>. Relevant discussion between analysts about how to proceed with the intended analysis or interpreting preliminary results are tracked within an issue’s thread. Each issue is assigned a unique number. Any code or analysis outputs, such as figures, tables or other files, should be committed and tagged with the issue number, so that all analysis addressing that task is tracked within the issue’s thread (Fig X, step 4, Do) <screen-cap?>. The URL for the GitHub issue should then be added to the relevant preregistration item, the preregistration document should then be committed with the changes, and the version number of the preregistration updated in a new release.

… And back again.

As the modelling proceeds, further detail on later phases of the analysis can be iteratively updated and preregistered. Where the results of model evaluation and analysis reveal that there are problems with the model, plans can be changed, and again the next phase can be preregistered. For example, if a researcher finds that assumptions are violated or other unexpected results occur that force a change to the planned analysis, the preregistration can be updated based on the findings of those analyses. When this occurs, the major version number of the preregistration document should be incremented (Fig X steps 5 - 6). Since the findings of the analyses are tracked within the relevant issue for that template item, and the issue is recorded in the preregistration itself, the trigger for the change in the plan is made explicitly clear.

For each altered or new preregistration item, a the issue thread should either be updated, or a new issue created respectively (Fig X, step 6). Those analyses are then either reconnected based on the revised plan or the new investigatory analyses are then conducted (Fig X, step 7).

This process continues until the preregistration is complete, and the researcher can continue to execute the plan as it has been fully described in the final version of the preregistration document (Fig X, step 8 - 9).

Flexible Strategies, Preliminary Analyses or Investigations

When a researcher needs to conduct preliminary analyses or check parts of the data before committing to a particular decision about the model or analysis, they can specify a flexible strategy in their response to that preregistration item by:

- stating what needs to be known to move forward with the modeling and analysis

- explaining how the researcher will test this, and what parts of the data will be used in that analysis

- describing how the results will be interpreted, listing each potential decision and the analysis result that will trigger that decision, where possible.

![Adaptive Preregistration Workflow - Research Workflow for implementing a ‘adaptive’ preregistration in a modelling context](../figures/LivingPreregistration.png)

--------------------------------

@@

*Determine the temporal and spatial scope of the model. Where is the boundary of the modelled system? Everything outside beyond the boundary and not crossing it is to be ignored within the domain of the model, and everything crossing the boundary is to be treated as external forcing (known/unknown), or else as model outputs (observed, or not, Jakeman et al. 2006) ***HF comment I dont understand the preceeding sentence, the concept of boundaries of the model is foreign to me. Also, is this question more to do with describing the scope and limitations of the data rather then the model?***. The choice of a model's boundaries is closely linked to the choice of how finely to aggregate the behaviour within the model (Jakeman et al., 2006) - what is the intended scale and resolution of the model (temporal, spatial or otherwise)?

@@

What is the intended domain of applicability of the model, what is the extent of acceptable extrapolations (Grimm)? This is relevant to model transferability... (similar to a COG statement..)* 
A comment:This section feels a bit too big and confusing from my perspective and I do think you might need to move some of it to a different section to reduce the risk of misinterpretation. I'm just thinking about how I might fill this out for an imagined study. The literal boundaries of the study are usually straightforward, but the theoretic boundaries are a different and also big thing to try to conceptualise in one section. You could have a section for the spatial/temporal boundaries of the study, and another section on how far the results can be extrapolated based on the study design. This makes sense to me in how I design experiments, if I have too many confounding variables and not enough spatial and environmental replication then I'm less likely to make any broad claims. However, if I have controlled or experimentally manipulated factors then I feel safer making bigger generalisations about results.

@@

* Describe how uncertainty and variation will be represented in this model. This includes both exogenous uncertainties affecting the system, parametric uncertainty in input data and structural / conceptual nonparametric uncertainty in the model [@Moallemi2019].*

@@

- *For each dataset, is the data open or publically available? (Y/N)*
- *How can the data be accessed? Provide a link or contact as appropriate, indicating any restrctions on the use of data.*
- *Date of download, access, or future access:*
- *Describe the source of the data - what entity originally collected this data? (National Data Set, Private Organisational Data, Own Lab Collection, Other Lab Collection, External Contractor, Meta-Analysis, Expert Elicitation, Other).*
- *Codebook and meta-data. If a Codebook or other meta-data is available, please link to it here and / or upload the document(s).*
- *Prior work based on this dataset - Have you published / presented any previous work based on this dataset? Include any publications, conference presentations (papers, posters), or working papers (in-prep, unpublished, preprints) based ont this dataset you have worked on.* HF- Love this question
- *Unpublished Prior Research Activity - Describe any prior but unpublished research activity using these data. Be specific and transparent.*
- *Prior knowledge of the current dataset - Describe any prior knowledge of or interaction with the dataset before commencing this study. For example, have you read any reports or publications about this data? *

@@

*Describe how you will determine how reliable the data is for the given purpose. Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009].* 
### 2.4.5 Data Evaluation: Data reliability reporting
### 2.4.5. Data Evaluation: data reliability reporting

@@

### 3.4.2 Choice of parameter estimation technique	
HF- I think that it might be double dipping to include structure here as well as in 3.3
*Specify what technique you will use to estimate parameter values, and how you will supply non-parametric variables and/or data (e.g. distributed boundary conditions). For example, will you calibrate all variables simultaneously by optimising fit of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination (Jakeman, 2006). Identify which variables were parameterised directly, such as by expert elicitation or prior knowledge. Specify which algorithm(s) you will use for any data-driven parameter estimation, including supervised, or unsupervised machine learning, decision-tree, K-nearest neighbour or cluster algorithms [@Liu2018b].*

@@

*Describe how you will validate and check the calibration of the model. The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation)  [@Yates2018;@Barnard2019]. Justify your choice*

@@

*"Model Checking" goes by many names ("conditional verification", "quantitative verification", "model output verification" <cite>), and refers to a series of analyses that assess a model's performance in representing the system of interest [@Conn:2018hd].  Model checking therefore heps diagnose assumption violations, and reveals where a model might need to be altered to better represent the data, and therefore system [@Conn:2018hd]. Quantitative model checking diagnostics include goodness of fit, tests on residuals or errors, such as for heteroscedascity, cross-correlation, and autocorrelation [@Jakeman:2006ii].*

@@

# 5.0 Model Validation and Evaluation
HF- I think that this is covered in section 4 with descriptions of validation and checking. I'm not sure how this section is different.

@@

## 5.1 Model output corroboration	
*Ideally, model outputs or predictions are compared to independent data and patterns that were not used to develop, parameterise, or verify the model. Testing against a dataset of response and predictor variables that are sptially and/or temporally independent from the training dataset¬†minimises the risk of artificially inflating model performance measures [@Araujo2019].State whether you will corroborate the model outputs on external data, and please document any independent validation data in step <2.4>.It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building, please describe the relationship modllers have to model validation data, will independent datasets be known to any modeller or analyst involved in the model building process?Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system ), model validation is not always possible because empirical experiments are infeasible or model users are working on rapid-response time-frames ‚Äî hence, why ecologists often model in the first place (Grimm, et al. 2014. Independent predictions can instead be tested on submodels. Alternatively, patterns in model output that are robust and seem characteristic of the system can be identified and evaluated in consultation with the literature or by experts to judge how accurate the model's output is (Grimm et al.2014). If unable to evaluate the model outputs against independent data, explain why and explain what steps you will take to interrogate the model.*

@@

## 5.3 Evaluate model assumptions	
HF- I would put this up in section 4 and get rid of section 5
*Explain how you will demonstrate a lack of violation of or robustness to model assumptions. If you cannot perform quantitative tests, specify what theoretical justifications would justify a lack of violation of or robustness to model assumptions. If you cannot demonstrate or theoretically justify violation or robustness to assumptions, explain why not, and specify whether you will discuss assumption violations and their consequences for interpretation of model outputs.(Complete in follow-up iterations of the pre-registration, only if there are departures from assumptions as demonstrated in the planned tests above). The consequences of assumption violations on the interpretation of results should be assessed [@Araujo2019]. Specify how you will assess departures from assumptions. If assumption violations cannot be avoided, explain describe how you will explore their consequences.*

## 5.4 Model analysis	
HF- I would put 5.4 and it's sbsections up in section 4 and get rid of section 5. Also, as above, remove the paragraph under 5.4
*Uncertainty in models arises due to incomplete system understanding (which processes to include, or which interact), from imprecise, finite and sparese data measurements, and from uncertainty in input conditions and scenarios for model simulations or runs [@Jakeman:2006ii]. Non-technical uncertainties can also be introduced throughout the modellign process, such as uncertainties arising from issues in problem-framing, indeterminicies, and modeller / client values [@Jakeman:2006ii].The purpose of model analysis is to prevent blind trust in the model by understanding how model outputs have emerged, and to ‚Äòchallenge' the model by verifying whether the model is still believeable and fit for purpose if one or more parameters are changed [@Grimm:2014es]. Model analysis should increase understanding of the model behaviour by identifying which processes and process interactions explain characteristic behaviours of the model system. Model analysis typically consists of sensitivity analyses preceded by uncertainty analyses [@Saltelli2019], and a suite of other simulation or other computational experiments. The aim of such computational experiments is to increase understanding of the model behaviour by identifying which processes and process interactions explain characteristic behaviours of the model system [@Grimm:2014es].Uncertainty analyses and sensitivity analyses augment one another to draw conclusions about model uncertainty ‚Äî the overall model will be very uncertain if the parameters that are the most uncertain are the ones the model is most uncertain to., for instance.Because the results from a full suite of sensitivity analysis and uncertainty analysis can be difficult to interpret due to the number and complexity of causal relations examined [@Jakeman:2006ii], it is useful for the analyst to relate the choice of analysis to the modelling context, purpose and analytical objectives defined in the problem formulation phase, in tandem with any critical uncertainties that have emerged during model development and testing prior to this point.*

@@

*Uncertainty can arise from different modeling techniques, response data and predictor variables [@Araujo2019]. Uncertainty analyses characterise the uncertainty in model outputs, and identify how uncertainty in model parameters affects uncertainty in model output, but does not identify which model assumptions are driving this behaviour [@Grimm:2014es;@Saltelli2019]. Uncertainty analyses can include propagating known uncertainties through the model, or by investigating the effect of different model scenarios with different parameters and modelling technique combinations [@Araujo2019], for example. It also includes characterising the output distribution, for example "by constructing it empirically from model output data points". It could also include extracting summary statistics like the mean, median and variance from this distribution, and perhaps constructing confidence intervals on the mean [@Saltelli2019].Please describe how you will characterise model and data uncertainties, as well as how you plan on summarising the results of these experiments with graphical, tabular or other devices. Relate your choice of analysis to the context and purposes of the model described in the problem formulation phase. For instance ‚Äúdiscrepancies between model output and observed output may be important for forecasting models, where cost, benefit, an risk over a substantial period must be gauged, but much less critical for decision-making or management models where the user may be satisfied with knowing that the predicted ranking order of impcats of alternative scenarios or management options is likely to be correct, with only a rough indication of their sizes" [@Jakeman:2006ii].If the chosen modelling approach is able to explicitly articulate uncertainty due to data, measurements or baseline conditions, such as by providing estimates of uncertainty (typically in the form of probabilistic paraameter covariance, [@Jakeman:2006ii]), specify which measure of uncertainty you will measure.*

@@

# 6.0 Model Application
HF- I feel that this belongs at the end of section 1
*Explain how you will operationalise any scenarios identified during the problem formulation phase, if not already done so.¬†For example, how will you operationalise any qualitative changes of interest, such as ‚Äòdeterioration' or ‚Äòimprovement'?Specify any input conditions and relevant parameter values for initial environmental conditions and decision-variables under each scenario. Please describe any other relevant technical details of model application, such as methods for how you will implement any simulations or model projections. Specify what raw and transformed model outputs will you extract from the model simulations or projections.*

## 6.1 Choose performance criteria for scenario analysis
HF- I feel that this belongs at the end of section 1
*Describe how you will evaluate and distinguish the performance of alternative scenario outcomes. These measures should link back to the analytical objectives specified during the problem formulation phase, and might include for example, descriptive statistic measures, such as a model in the top 10th percentile of worst conditions, satisficing measures such as a minimum performance threshold for achieving some management objective, or perhaps regret measures, such as choosing a decision with minimum regret [@Moallemi2019]."Justify or otherwise explain how you chose these measures, and determined performance criteria in relation to the analytical objectives, modelling context, such as the risk attitudes of decision-makers and stakeholders within this system.*

## 6.2 Synthesis of model application results	
*Specify how you will map, plot or otherwise display and synthesise the results of scenario and model analyses. Explain how will they support meaningful analysis and interpretation to aid in answering the analytical objectives and modelling context specified in the problem formulation phase.Explain how you will analyse the model analysis output to answer your analytical objectives. For instance, describe any trade-off or robustness analyses you will undertake to help evaluate and choose between different alternatives in consultation with experts or decision-makers.*

## 6.3 Consultation with Experts, Clients, Stakeholders	
HF- I'm not sure how possible people will find doing this... and whether it is potentially problematic to ask people to because their clients will amost certainly change their mind about how they deal with the information at this point in a way that the researcher cant control. 
*How will you use the results of model application analysis to provide advice. and support to the clients / managers / stakeholders based on these findings? Describe how these results will be characterised and presented to experts, clients and stakeholders once the results from the modelling exercises have been synthesised. What role will these consultations have in determining the selection of management interventions, or future iterations of model development? For example, what key outliers or strongest examples of best or worst case scenarios should be presented to experts or decision-makers?(Do we need an item about how do we interpret the model outputs? Should there be some a priori decisions about how we interpret the model findings, and how does case 49 about the risk attitude inform how the analyst interprets the model findings?)*

>>>>

@@

*Determine the temporal and spatial scope of the model. Where is the boundary of the modelled system? Everything outside beyond the boundary and not crossing it is to be ignored within the domain of the model, and everything crossing the boundary is to be treated as external forcing (known/unknown), or else as model outputs (observed, or not, Jakeman et al. 2006)* **HF comment I dont understand the preceeding sentence, the concept of boundaries of the model is foreign to me. Also, is this question more to do with describing the scope and limitations of the data rather then the model?**\*.
The choice of a model's boundaries is closely linked to the choice of how finely to aggregate the behaviour within the model (Jakeman et al., 2006) - what is the intended scale and resolution of the model (temporal, spatial or otherwise)?

@@

What is the intended domain of applicability of the model, what is the extent of acceptable extrapolations (Grimm)?
This is relevant to model transferability... (similar to a COG statement..)\* A comment:This section feels a bit too big and confusing from my perspective and I do think you might need to move some of it to a different section to reduce the risk of misinterpretation.
I'm just thinking about how I might fill this out for an imagined study.
The literal boundaries of the study are usually straightforward, but the theoretic boundaries are a different and also big thing to try to conceptualise in one section.
You could have a section for the spatial/temporal boundaries of the study, and another section on how far the results can be extrapolated based on the study design.
This makes sense to me in how I design experiments, if I have too many confounding variables and not enough spatial and environmental replication then I'm less likely to make any broad claims.
However, if I have controlled or experimentally manipulated factors then I feel safer making bigger generalisations about results.

@@

-   Describe how uncertainty and variation will be represented in this model. This includes both exogenous uncertainties affecting the system, parametric uncertainty in input data and structural / conceptual nonparametric uncertainty in the model [@Moallemi2019].\*

@@

-   *For each dataset, is the data open or publically available? (Y/N)*
-   *How can the data be accessed? Provide a link or contact as appropriate, indicating any restrctions on the use of data.*
-   *Date of download, access, or future access:*
-   *Describe the source of the data - what entity originally collected this data? (National Data Set, Private Organisational Data, Own Lab Collection, Other Lab Collection, External Contractor, Meta-Analysis, Expert Elicitation, Other).*
-   *Codebook and meta-data. If a Codebook or other meta-data is available, please link to it here and / or upload the document(s).*
-   *Prior work based on this dataset - Have you published / presented any previous work based on this dataset? Include any publications, conference presentations (papers, posters), or working papers (in-prep, unpublished, preprints) based ont this dataset you have worked on.* HF- Love this question
-   *Unpublished Prior Research Activity - Describe any prior but unpublished research activity using these data. Be specific and transparent.*
-   *Prior knowledge of the current dataset - Describe any prior knowledge of or interaction with the dataset before commencing this study. For example, have you read any reports or publications about this data?*

@@

*Describe how you will determine how reliable the data is for the given purpose. Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009].* \### 2.4.5 Data Evaluation: Data reliability reporting \### 2.4.5.
Data Evaluation: data reliability reporting

@@

### 3.4.2 Choice of parameter estimation technique

HF- I think that it might be double dipping to include structure here as well as in 3.3 *Specify what technique you will use to estimate parameter values, and how you will supply non-parametric variables and/or data (e.g. distributed boundary conditions). For example, will you calibrate all variables simultaneously by optimising fit of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination (Jakeman, 2006). Identify which variables were parameterised directly, such as by expert elicitation or prior knowledge. Specify which algorithm(s) you will use for any data-driven parameter estimation, including supervised, or unsupervised machine learning, decision-tree, K-nearest neighbour or cluster algorithms [@Liu2018b].*

@@

*Describe how you will validate and check the calibration of the model. The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation) [@Yates2018; @Barnard2019]. Justify your choice*

@@

*"Model Checking" goes by many names ("conditional verification", "quantitative verification", "model output verification"* <cite>), and refers to a series of analyses that assess a model's performance in representing the system of interest [@Conn:2018hd].
Model checking therefore heps diagnose assumption violations, and reveals where a model might need to be altered to better represent the data, and therefore system [@Conn:2018hd].
Quantitative model checking diagnostics include goodness of fit, tests on residuals or errors, such as for heteroscedascity, cross-correlation, and autocorrelation [@Jakeman:2006ii].

@@

# 5.0 Model Validation and Evaluation

HF- I think that this is covered in section 4 with descriptions of validation and checking.
I'm not sure how this section is different.

@@

## 5.1 Model output corroboration

*Ideally, model outputs or predictions are compared to independent data and patterns that were not used to develop, parameterise, or verify the model. Testing against a dataset of response and predictor variables that are sptially and/or temporally independent from the training dataset¬†minimises the risk of artificially inflating model performance measures [@Araujo2019].State whether you will corroborate the model outputs on external data, and please document any independent validation data in step \<2.4\>.It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building, please describe the relationship modllers have to model validation data, will independent datasets be known to any modeller or analyst involved in the model building process?Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system ), model validation is not always possible because empirical experiments are infeasible or model users are working on rapid-response time-frames ‚Äî hence, why ecologists often model in the first place (Grimm, et al. 2014. Independent predictions can instead be tested on submodels. Alternatively, patterns in model output that are robust and seem characteristic of the system can be identified and evaluated in consultation with the literature or by experts to judge how accurate the model's output is (Grimm et al.2014). If unable to evaluate the model outputs against independent data, explain why and explain what steps you will take to interrogate the model.*

@@

## 5.3 Evaluate model assumptions

HF- I would put this up in section 4 and get rid of section 5 *Explain how you will demonstrate a lack of violation of or robustness to model assumptions. If you cannot perform quantitative tests, specify what theoretical justifications would justify a lack of violation of or robustness to model assumptions. If you cannot demonstrate or theoretically justify violation or robustness to assumptions, explain why not, and specify whether you will discuss assumption violations and their consequences for interpretation of model outputs.(Complete in follow-up iterations of the pre-registration, only if there are departures from assumptions as demonstrated in the planned tests above). The consequences of assumption violations on the interpretation of results should be assessed [@Araujo2019]. Specify how you will assess departures from assumptions. If assumption violations cannot be avoided, explain describe how you will explore their consequences.*

## 5.4 Model analysis

HF- I would put 5.4 and it's sbsections up in section 4 and get rid of section 5.
Also, as above, remove the paragraph under 5.4 *Uncertainty in models arises due to incomplete system understanding (which processes to include, or which interact), from imprecise, finite and sparese data measurements, and from uncertainty in input conditions and scenarios for model simulations or runs [@Jakeman:2006ii]. Non-technical uncertainties can also be introduced throughout the modellign process, such as uncertainties arising from issues in problem-framing, indeterminicies, and modeller / client values [@Jakeman:2006ii].The purpose of model analysis is to prevent blind trust in the model by understanding how model outputs have emerged, and to ‚Äòchallenge' the model by verifying whether the model is still believeable and fit for purpose if one or more parameters are changed [@Grimm:2014es]. Model analysis should increase understanding of the model behaviour by identifying which processes and process interactions explain characteristic behaviours of the model system. Model analysis typically consists of sensitivity analyses preceded by uncertainty analyses [@Saltelli2019], and a suite of other simulation or other computational experiments. The aim of such computational experiments is to increase understanding of the model behaviour by identifying which processes and process interactions explain characteristic behaviours of the model system [@Grimm:2014es].Uncertainty analyses and sensitivity analyses augment one another to draw conclusions about model uncertainty ‚Äî the overall model will be very uncertain if the parameters that are the most uncertain are the ones the model is most uncertain to., for instance.Because the results from a full suite of sensitivity analysis and uncertainty analysis can be difficult to interpret due to the number and complexity of causal relations examined [@Jakeman:2006ii], it is useful for the analyst to relate the choice of analysis to the modelling context, purpose and analytical objectives defined in the problem formulation phase, in tandem with any critical uncertainties that have emerged during model development and testing prior to this point.*

@@

*Uncertainty can arise from different modeling techniques, response data and predictor variables [@Araujo2019]. Uncertainty analyses characterise the uncertainty in model outputs, and identify how uncertainty in model parameters affects uncertainty in model output, but does not identify which model assumptions are driving this behaviour [@Grimm:2014es; @Saltelli2019]. Uncertainty analyses can include propagating known uncertainties through the model, or by investigating the effect of different model scenarios with different parameters and modelling technique combinations [@Araujo2019], for example. It also includes characterising the output distribution, for example "by constructing it empirically from model output data points". It could also include extracting summary statistics like the mean, median and variance from this distribution, and perhaps constructing confidence intervals on the mean [@Saltelli2019].Please describe how you will characterise model and data uncertainties, as well as how you plan on summarising the results of these experiments with graphical, tabular or other devices. Relate your choice of analysis to the context and purposes of the model described in the problem formulation phase. For instance ‚Äúdiscrepancies between model output and observed output may be important for forecasting models, where cost, benefit, an risk over a substantial period must be gauged, but much less critical for decision-making or management models where the user may be satisfied with knowing that the predicted ranking order of impcats of alternative scenarios or management options is likely to be correct, with only a rough indication of their sizes" [@Jakeman:2006ii].If the chosen modelling approach is able to explicitly articulate uncertainty due to data, measurements or baseline conditions, such as by providing estimates of uncertainty (typically in the form of probabilistic paraameter covariance, [@Jakeman:2006ii]), specify which measure of uncertainty you will measure.*

@@

# 6.0 Model Application

HF- I feel that this belongs at the end of section 1 *Explain how you will operationalise any scenarios identified during the problem formulation phase, if not already done so.¬†For example, how will you operationalise any qualitative changes of interest, such as ‚Äòdeterioration' or ‚Äòimprovement'?Specify any input conditions and relevant parameter values for initial environmental conditions and decision-variables under each scenario. Please describe any other relevant technical details of model application, such as methods for how you will implement any simulations or model projections. Specify what raw and transformed model outputs will you extract from the model simulations or projections.*

## 6.1 Choose performance criteria for scenario analysis

HF- I feel that this belongs at the end of section 1 *Describe how you will evaluate and distinguish the performance of alternative scenario outcomes. These measures should link back to the analytical objectives specified during the problem formulation phase, and might include for example, descriptive statistic measures, such as a model in the top 10th percentile of worst conditions, satisficing measures such as a minimum performance threshold for achieving some management objective, or perhaps regret measures, such as choosing a decision with minimum regret [@Moallemi2019]."Justify or otherwise explain how you chose these measures, and determined performance criteria in relation to the analytical objectives, modelling context, such as the risk attitudes of decision-makers and stakeholders within this system.*

## 6.2 Synthesis of model application results

*Specify how you will map, plot or otherwise display and synthesise the results of scenario and model analyses. Explain how will they support meaningful analysis and interpretation to aid in answering the analytical objectives and modelling context specified in the problem formulation phase.Explain how you will analyse the model analysis output to answer your analytical objectives. For instance, describe any trade-off or robustness analyses you will undertake to help evaluate and choose between different alternatives in consultation with experts or decision-makers.*

## 6.3 Consultation with Experts, Clients, Stakeholders

HF- I'm not sure how possible people will find doing this... and whether it is potentially problematic to ask people to because their clients will amost certainly change their mind about how they deal with the information at this point in a way that the researcher cant control.
*How will you use the results of model application analysis to provide advice. and support to the clients / managers / stakeholders based on these findings? Describe how these results will be characterised and presented to experts, clients and stakeholders once the results from the modelling exercises have been synthesised. What role will these consultations have in determining the selection of management interventions, or future iterations of model development? For example, what key outliers or strongest examples of best or worst case scenarios should be presented to experts or decision-makers?(Do we need an item about how do we interpret the model outputs? Should there be some a priori decisions about how we interpret the model findings, and how does case 49 about the risk attitude inform how the analyst interprets the model findings?)*
